
\section{Gcylc Config File Reference}
\label{GcylcRCReference}

\lstset{language=bash}

This section defines all legal items and values for the gcylc user
config file, which should be located in
\lstinline=$HOME/.cylc/gcylc.rc=.
Current settings can be printed with the \lstinline=cylc get-gui-config=
command.

\subsection{Top Level Items}

\subsubsection{initial views}

Set the suite view panel(s) displayed initially, when the GUI starts.
This can be changed later using the tool bar.

\begin{myitemize}
\item {\em type:} string (a list of one or two view names)
\item {\em legal values:} ``text'', ``dot'',  ``graph''
\item {\em default:} ``text''
\item {\em example:} \lstinline@initial views = graph, dot@
\end{myitemize}

\subsubsection{ungrouped views}

List suite views, if any, that should be displayed initially in an
ungrouped state. Namespace family grouping can be changed later
using the tool bar.

\begin{myitemize}
\item {\em type:} string (a list of zero or more view names)
\item {\em legal values:} ``text'', ``dot'',  ``graph''
\item {\em default:} (none)
\item {\em example:} \lstinline@ungrouped views = text, dot@
\end{myitemize}

\subsubsection{use theme}

Set the task state color theme, common to all views, to use
initially. The color theme can be changed later using the tool bar. 
See \lstinline@gcylc.rc.eg@ and \lstinline@themes.rc@ in
\lstinline@$CYLC_DIR/conf/gcylcrc/@ for how to modify existing color
themes or define your own. Use \lstinline@cylc get-gui-config@
to list your available themes.

\begin{myitemize}
\item {\em type:} string (theme name)
\item {\em legal values:} ``default'', ``solid'', ``high-contrast'',
    ``color-blind'', and any custom or user-modified themes.
\item {\em default:} ``default''
\end{myitemize}

\subsubsection{dot icon size}

Set the size of the task state dot icons displayed in the text and dot
views.

\begin{myitemize}
\item {\em type:} string
\item {\em legal values:} ``small'' (10px), ``medium'' (14px), ``large'' (20px)
\item {\em default:} ``medium''
\end{myitemize}

\subsubsection{sort by definition order}

If this feature is turned on the default sort order for task names and
families in the dot and text views will the order that the appear in the
suite definition. Clicking on the task name column in the treeview will
toggle to alphanumeric sort, and a View menu item does the same for the
dot view.  If turned off, the default sort order is alphanumeric and
definition order is not available at all.

\begin{myitemize}
\item {\em type:} boolean
\item {\em default:} False
\end{myitemize}

\subsection{[themes]}

This section may contain task state color theme definitions.

\subsubsection[THEME]{[themes] $\rightarrow$ THEME}

The name of the task state color-theme to be defined in this section.

\begin{myitemize}
\item {\em type:} string
\end{myitemize}

\subsubsection[inherit]{[themes] $\rightarrow$ [THEME] $\rightarrow$ inherit}

You can inherit from another theme in order to avoid defining all states.

\begin{myitemize}
\item {\em type:} string (parent theme name)
\item {\em default:} ``default''
\end{myitemize}

\subsubsection[defaults]{[themes] $\rightarrow$ [THEME] $\rightarrow$ defaults}

Set default icon attributes for all state icons in this theme.

\begin{myitemize}
\item {\em type:} string list (icon attributes)
\item {\em legal values:} ``color=COLOR``, ``style=STYLE'', ``fontcolor=FONTCOLOR''
\item {\em default:} (none)
\end{myitemize}

For the attribute values, COLOR and FONTCOLOR are X color names or hex codes (e.g.
\lstinline@#ff0000@ for red); and STYLE can be ``filled'' or ``unfilled''. 
See \lstinline@gcylc.rc.eg@ and \lstinline@themes.rc@ in
\lstinline@$CYLC_DIR/conf/gcylcrc/@ for examples. 

\subsubsection[STATE]{[themes] $\rightarrow$ [THEME] $\rightarrow$ STATE}

Set icon attributes for all task states in THEME, or for a subset of them if 
you have used theme inheritance and/or defaults. Legal values of STATE are
any of the cylc task proxy states: {\em waiting, runahead, held, queued, ready, 
submitted, submit-failed, running, succeeded, failed, retrying, submit-retrying}.
 
\begin{myitemize}
\item {\em type:} string list (icon attributes)
\item {\em legal values:} ``color=COLOR``, ``style=STYLE'', ``fontcolor=FONTCOLOR''
\item {\em default:} (none)
\end{myitemize}

For the attribute values, COLOR and FONTCOLOR are X color names or hex codes (e.g.
\lstinline@#ff0000@ for red); and STYLE can be ``filled'' or ``unfilled''. 
See \lstinline@gcylc.rc.eg@ and \lstinline@themes.rc@ in
\lstinline@$CYLC_DIR/conf/gcylcrc/@ for examples. 

